Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture

### 2013-10-17 ieee_paper.tex
I am confused. I possibly screwd things up.

\section{Introduction}
The multiplication of sparse matrix by a dense vector (SpMV) is an
important kernel in many fields from scientific computing~\cite{} \todo{Do you have a good citation for that?}
to graph analysis (recommender systems~\cite{Brin98}, graph
drawing~\cite{Koren05}). So improving the performance of SpMV has
captured the interest of many researcher; including the development of
various implementations for multi-core
CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
  conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
challenge that is faced to obtain good performance for SpMV is that
the operation are conducted using memory location that are irregular
and often unpredictible. That makes the kernels mostly memory
bound and there is a significant instruction overhead per flop.

Common improvement techniques such as register blocking, bandwidth
reduction (matrix reordering~\cite{Cuthill69}), partitioning to fit in
cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
unrolling~\cite{Mellor-Crummey04} have impacts which are very
dependent on the matrix and overall do not lead to dramatic
improvement. (The state-of-the-art techniques in OSKI~\cite{Vuduc05}
provide some useful yet limited improvements). Register blocking~\cite{conf/ppsc/Toledo97} does
not apply well to most the matrix at hand (despite it can be applied
with virtually no overhead in all cases thanks to compressed
representation~\cite{Buluc11}). There is about 8 bytes of the matrix to move
in per non zero (assuming single precision); each nonzero requires two
floating point operations leading to a flop-to-byte ratio of at most
$\frac{1}{4}$. This limits the obtained performance to at most a
quarter of the bandwidth of the architecture wasting a lot of
potentially useful cycles. The commonly used techniques are mostly
designed to reach that bound rather than overcome it.

Fortunately that fate is not inevitable. One solution would be to
schedule a more instruction intensive kernel simultaneously with the
execution of SpMV, relying on some hardware threading capabilities,
such as HyperThreading, to reduce the cycle wastage. However, most ot
the applications that use SpMV do not typically have an instruction
intensive kernel to run simultaneously.

An other solution, and the one we pursue in this paper, is to compute
multiple SpMV at once on matrices that have the same sparsity
patterns. Obviously not all the applications have such a property. But
important classes of applications such as graph
recommendation~\cite{Kucuktunc13-SNAM}, eigensolving~\cite{LOBPCG} and the computation of
derivative of Radial Basis Functions(RBF)~\cite{} \todo{a good citation?} can use multiple
SpMVs simultaneously. In this paper in particular, we investigate the
case of the derivative of RBFs where four derivatives of four
different functions is expressed as the multiplication of four vectors
by four matrices with identical sparsity patterns leading to the
simultaneous execution of 16 SpMVs at a time. Using multiple vectors
at a time has been investigated before to densify the
computations~\cite{Im01} but this is the first time that the
densification comes from adding both vectors and matrices.

@@

In Section~\ref{sec:rbf} we present the computation of RBFs and how it
can be expressed 16 multiplication of 4 vectors by 4 sparse matrices
with a common sparsity pattern. Section~\ref{sec:model} presents an
estimation of the instruction intensity of various form of the
computations and we show that a 6- fold improvement can be
expected when computing the 16 multiplications simultaneously and
reach a total of about 210 Gflop/s. This performance represents
approximatively 10\% of the available flop/s of a Knight Corner
coprocessor. Therefore, it is necessary to have implementations that
perform the computation in as little amount of instructions as
possible. We describe in Section~\ref{sec:impl} the details of the MIC
architecture and how to use specialized load, store, swizzle and
permutation instruction to efficiently bring the data in the vector
registers to be processed. Section~\ref{sec:expe} gives some
experimental result about the amount of bandwidth that can be achieved
depending on how the SpMV kernel is written and the actual performance
of the various kernel on multiple classes of matrices some generated
for analysis purpose and some extracted from and RBF application. A
performance of 135 Gflop/s is achieved on real scenario. Concluding
remarks and perspectives are provided in Section~\ref{sec:ccl}.

@@

where $f$ is the source vector of function values and $g$ is the
resulting vector of derivative values. The computation of a single
derivative has been reduced to a SpMV, where each row has $n_z$
nonzeros. In
practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
three-dimensional flows. These numbers are similar to what is used in
finite-element codes. Notice that the non zero elements of the matrix
correspond to a relation of nearest neighbor. This relationship is not
necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
means that the matrix sparsity pattern might not be symmetric either.

@@

\todo{describe the data structure} Since the sparsity of each row is
constant, ELLPACK becomes the most appropriate matrix
compression scheme.

@@

\begin{figure}[bth]
  \begin{center}
    \scalebox{.9}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of non-zero\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the matrice(s) in bytes\\
        $s_x$ & size of the {\tt x} vector(s) in bytes\\
        $s_y$ & size of the {\tt y} vector(s) in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred  \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred  \\
        $O$     & number of floating point operations \\
        $I_b$   & maximal computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notations}
  \label{tab:not}
\end{figure}

Each vector in the problem is of dimension $n_r$ and each entry
takes $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
  y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
$$s_x = n_v b_x n_r$$ $$s_y = n_v n_m b_x n_r$$

@@

If we assume an algorithm where the rows are processed the one after
the other, the amount of memory written is precisely the
size of the {\tt y} vector. (This assumption removes the possibility of
blocking or cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$

The amount of data read from memory depends highly on both the
algorithm's execution path, and on how the matrix is structured. But
in the best case both the matrix $A$ and the source vector {\tt x} are
read once from the main memory. (We assume that all the elements of
$\tt x$ are involved in the SpMV.)  Thus, $$b_{rT} = s_M + s_x = n_r
n_z (b_i + b_x n_m) + n_v b_x n_r$$
 $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$

Notice that there is no reason for a piece of the matrix to be read
multiple times. But assuming that each element of the {\tt x} vector
is read a single time is a strong assumption. If using a single core,
it assumes that either the cache of the architecture can store the full
{\tt x} vector or that the matrix is sufficiently well structured 
to cause no cache trashing. If using multiple cores, this assumes that
no element of the {\tt x} vectors will be used by multiple
cores. \cite{Saule13-ARXIV} showed that there is very little cache trashing
in practice, but it showed that having elements of the vectors used by
multiple cores can have a significant impact on the performance
(growing with $n_v$).

On the other hand, in the worst case, every time the {\tt x} vector is
accessed, the value needs to be transfered from memory again. So in
total, there are as many transfers as the number of non-zeros in the
matrix. Note however that most architectures cannot read memory a
single byte at a time. Instead, a minimum number of bytes, equal to
the size of a cacheline $cl$, are transfered at once.  When there are
multiple vectors, each non-zero element uses $n_v$ consecutive
entries. The worst case number of bytes read and transfered are
$$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
$$B_T = n_v n_m b_x n_r + n_r n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)$$

@@

%% In the special case $n_v=n_m=1$ in single precision
%% $$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
%% %$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
%% $$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$

%% In the special case $n_v=n_m=4$ in single precision
%% $$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
%% %$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
%% $$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
%% The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.

%% \begin{table}
%%   \centering
%%   \begin{tabular}{|l|r|r|r|r|}
%%     \hline
%%                            & v1m1  & v1m4  & v4m1   & v4m4  \\
%%     \hline
%%     Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
%%     Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
%%     Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
%%     Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
%%     \hline
%%   \end{tabular}
%%   \caption{Estimation of performance in GFlop/s at 150GB/s for a
%%     matrix of $n_z = 32$}
%% \end{table}

Figure~\ref{fig:ratio_bytes_flops} presents the flop to byte ratios
for single precision computation in the best and the worst case on a
classical cache based architecture ($cl = 64$) and assuming $cl=1$. We
can easily see the potential improvement in the computational
intensity when the number of matrices of vectors increase. There is a
significant difference between the best and the worst case: there is a
$8$ fold difference in the 1 matrix 1 vector case but that gap closes
with the increase in the number of vectors and matrices to a 4 fold
difference in the 4 matrices and 4 vectors case and 1.6 fold in the 16
matrices and 16 vectors. The ratios computed with $cl = 1$ are mostly
similar to the the one with $cl=64$ when the number of vectors is
large. But the differences are important when the number of vector is
low: this highlight that accessing the memory per batch is the main
problem faced when performing a standard SpMV computation.

\begin{figure*}
  \centering
  \subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}\label{fig:ratio_best}}%
  \subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}\label{fig:ratio_worst}}%
  \subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}\label{fig:ratio_worst_nocache}}

  \caption{Ratio of flops to bytes in single precision: in the best case in
    \ref{fig:ratio_best} best; in the worst case when $cl=64$ in
    \ref{fig:ratio_worst}; and in the worst case neglecting that the
    memory transfers are by batch (equivalent to $cl = 1$) in
    \ref{fig:ratio_worst_nocache}.}
  \label{fig:ratio_bytes_flops}
\end{figure*}

Figure~\ref{fig:ratio_bytes_flops} shows how these ratios translate
into actual performance. This figure provides projected best and worst
Gflop/s achievable assuming the computation is memory bound and the
architecture reaches 150Gflop/s. Notice that \cite{Saule13-ARXIV}
showed a higher peak bandwidth, but we will show in
Section~\ref{sec:expe} why 150Gflop/s is a better estimate of what one
might achieve in these kernels. In single precision, the worse that
one could achieve by using 4 vectors and 4 matrices is much higher
than the best achievable using a classical SpMV computation. The best
performance reachable is 210Gflop/s: almost 6 times higher than the
peak of the classical SpMV case.

\begin{figure*}
  \centering 
  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
%
  \subfigure[Relative performance to using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{Estimation of the reachable performance one can achieve
    using a device with a memory to computational units of 150GB/s
    when varying the number of vectors and matrices. \todo{we could
      remove the speedup plot if we need space.}  \todo{can the legend
      read "Single Precision (best)"} \todo{there is a missing
      parenthesis in the title; alternatively remove title.}}
  \label{fig:perf_predict}
\end{figure*}

@@

\begin{figure*}
\centering
  \subfigure[
    Entries with "cpp" denote cases where
    coding was performed without MIC vector
    instructions.]{\includegraphics[width=.40\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
%
  \subfigure[ xxx ] 
  {\includegraphics[width=.40\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}} \\

  \subfigure[The greater
    speed of the cpp version is obtained throught the use of
    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
    is aligned on 64
    bytes.]{\includegraphics[width=.40\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
%
  \subfigure[
    The speed on the CPU matches the speed with vector instructions.
    All memory is aligned on 32
    bytes.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}

@@

@@

\begin{figure*}
  \centering
  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
    solid circles: 4/4 case implemented in C++, solid triangles: 4/4
    case implemented with MIC vector instructions. $-O3$ compilation
    options. Each symbol has two cases: with (red) and without blue) 
    with and without Reverse Cuthill-McKee. The grid is $96^3$.]
    {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
%
  \subfigure[Performance on the host on a sparse matrix with $64^3$ rows. Manual
    vectorization brings significant improvement there as
    well.]{\includegraphics[width=.40\linewidth]{figures/64x64x64.pdf}\label{fig:perf_host}}
  \caption{Vectorization matters in practice. Differences in BW
    translate in different performance.\todo{this two plots should ``correlate'' }}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}

  \caption{Performance on a MIC of derivative computation for RBFFD
    with 884,736 nodes distributed quasi-randomly in a cube. Single
    derivative of a single function (base case, triangle) and four
    derivatives of four functions (multi case, circles). No bandwidth
    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
\end{figure}

>>>>

\section{Introduction}
\cite{Stock12-TACC} is all dense. \cite{
  Molka:2009:MPC:1636712.1637764} is about memory modeling not
directly spmv.

\cite{DBLP:journals/ijhpca/ShantharamCR11,
  DBLP:journals/corr/abs-1101-0091 }

The multiplication of sparse matrix by a dense vector (SpMV) is an
important kernel in many fields from scientific computing~\cite{} to
graph analysis (recommender systems~\cite{Brin98}, graph
drawing~\cite{Koren05}). So improving the performance of SpMV has
captured the interest of many researcher; including the development of
various implementations for multi-core
CPUs~\cite{Buluc2009SPAA,Williams07} and
GPUs~\cite{Bell08, conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
challenge that is faced to obtain good performance for SpMV is that
the operation are conducted using memory location that are irregular
and often unpredictible. That make that the kernels are mostly memory
bound and there is a significant instruction overhead per flop.

Common improvement techniques such as register blocking, bandwidth
reduction (matrix reordering~\cite{Cuthill69}), partitioning to fit in
cache or TLB~\cite{Nishtala07,Temam:1992:CBS:147877.148091,conf/ppsc/Toledo97},
unrolling~\cite{Mellor-Crummey04} have impacts which are very
dependent on the matrix and overall do not lead to dramatic
improvement. (The state-of-the-art techniques in OSKI~\cite{Vuduc05}
provide some useful yet limited improvements). Register blocking~\cite{conf/ppsc/Toledo97} does
not apply well to most the matrix at hand (despite it can be applied
with virtually no overhead in all cases thanks to compressed
representation~\cite{Buluc11}). there is about 8 bytes of the matrix to move
in per non zero (assuming single precision); each nonzero requires two
floating point operations leading to a flop-to-byte ratio of at most
$\frac{1}{4}$. This limits the obtained performance to at most a
quarter of the bandwidth of the architecture wasting a lot of
potentially useful cycles. The commonly used techniques are mostly
designed to reach that bound rather than overcome it.

Fortunately that fate is not inevitable. One solution would be to pair
multiple component of an application to schedule a more instruction
intensive kernel simultaneously with SpMV, relying on some hardware
threading capabilities, such as HyperThreading, to reduce the cycle
wastage. However most ot the applications that use SpMV do not
typically have an instruction intensive kernel to run simultaneously.

An other solution, and the one we pursue in this paper, is to compute
multiple SpMV at once on matrices that have the same sparsity
patterns. Obviously not all the applications have such a property. But
important classes of applications such as graph
recommendation~\cite{}, eigensolving~\cite{} and the computation of
derivative of Radial Basis Functions(RBF)~\cite{} can use multiple
SpMVs simultaneously. In this paper in particular, we investigate the
case of the derivative of RBFs where four derivatives of four
different function is expressed as the multiplication of four vectors
by four matrices with identical sparsity patterns leading to the
simultaneous execution of 16 SpMVs at a time. Using multiple vectors
at a time has been investigated before to densify the
computations~\cite{Im01} but this is the first time that the
densification comes from adding both vectors and matrices.

@@

In Section~\ref{sec:rbf} we present the computation of RBFs and how it
can be expressed 16 multiplication of 4 vectors by 4 sparse matrices
with a common sparsity pattern. Section~\ref{sec:model} presents an
estimation of the instruction intensity of various form of the
computations and we show that a 5- to 6- fold improvement can be
expected when computing the 16 multiplications simultaneously and
reach a total of about 200 Gflop/s. This performance represents
approximatively 10\% of the available flop/s of a Knight Corner
coprocessor. Therefore, it is necessary to have implementations that
perform the computation in as little amount of instructions as
possible. We describe in Section~\ref{sec:impl} the details of the MIC
architecture and how to use specialized load, store, swizzle and
permutation instruction to efficiently bring the data in the vector
registers to be processed. Section~\ref{sec:expe} gives some
experimental result about the amount of bandwidth that can be achieved
depending on how the spmv kernel is written and the actual performance
of the various kernel on multiple classes of matrices some generated
for analysis purpose and some extracted from and RBF application. A
performance of xxx GFlop/s is achieved on real scenario. Concluding
remarks and perspectives are provided in Section~\ref{sec:ccl}.

@@

where $f$ is the source vector of function values and $g$ is the
resulting vector of derivative values. The computation of a single
derivative has been reduced to a SpMV, where each row has $n_z$
nonzeros. Since the sparsity of each row is constant, ELLPACK~\cite{}
becomes the most appropriate matrix compression scheme\cite{}. In
practice, $n_z=32$ in two-dimensional flows and 64 or 100 for
three-dimensional flows. These numbers are similar to what is used in
finite-element codes. Notice that the non zero elements of the matrix
correspond to a relation of nearest neighbor. This relationship is not
necessarily symmetric as shown in Figure~\ref{fig:rbf_stencils} which
means that the matrix sparsity pattern might not be symmetric either.

@@

@@

\begin{figure}
  \begin{center}
    \scalebox{.9}{
      \begin{tabular}{|c|l|}
        %\hline
        %& & \\
        \hline
        $b_i$ & number of bytes per index \\
        $b_x$ & number of bytes per value \\
        $n_z$ & number of nonzeros per row $A$ \\
        $n_r$ & number of column/rows of $A$ \\
        $n_c$ & total number of non-zero\\
        $n_v$ & number of {\tt x} vectors \\
        $n_m$ & number of matrices \\
        $s_M$ & size of the matrice(s) in bytes\\
        $s_x$ & size of the {\tt x} vector(s) in bytes\\
        $s_y$ & size of the {\tt y} vector(s) in bytes\\
        \hline
        $cl$    & size of a cache line in bytes\\
        $b_{wT}$ & number of bytes written to memory  \\
        $b_{rT}$ & minimum number of bytes read from memory  \\
        $b_T$   & minimum number of bytes transferred  \\
        $B_{rT}$ & maximum number of bytes read from memory \\
        $B_T$   & maximum number of bytes transferred  \\
        $O$     & number of floating point operations \\
        $I_b$   & maximal computational intensity\\
        $I_w$   & minimum computational intensity\\        
        \hline
      \end{tabular}
    }
  \end{center}
  \caption{Notations}
  \label{tab:not}
\end{figure}

Each vector in the problem is of dimension $n_r$ and each dimension
tasks $b_x$ bytes. There are $n_v$ {\tt x} vectors and $n_v n_m$ {\tt
  y} vectors, which lead to the size of the {\tt x} and {\tt y} vectors:
$$s_x = n_v b_x n_r$$ $$s_y = n_v n_m b_x n_r$$

@@

If we assume an algorithm where the rows are processed one after
the other, the amount of memory written is precisely the
size of the {\tt y} vector. (This assumption removes the possibility of
cache partitioning techniques.) $$b_{wT} = s_y = n_v n_m b_x n_r$$

The amount of data read from memory depends highly on both the algorithm's
execution path, and on %algorithm execute, 
how the matrix is structured. But in the best case
both the matrix $A$ %is read a single time 
and the source vector {\tt x} are read once
%read a single time 
from the main memory. (We assume that all the elements of $\tt x$ are involved
in the SpMV.)  %is used is reasonnable since only oddly shaped matrix will
%left a dimension of the vector unused.) 
Thus, $$b_{rT} = s_M + s_x = n_r n_z
(b_i + b_x n_m) + n_v b_x n_r$$
 $$b_T = b_{rT} + b_{wT} =  n_r n_z (b_i + b_x n_m) + n_v b_x n_r (1 + n_m)$$

Notice that there is no reason for a piece of the matrix to be read
multiple times. But assuming that each element of the {\tt x} vector
is read a single time is a strong assumption. If using a single core,
it assumes that either the cache of the architecture can store the full
{\tt x} vector or that the matrix is sufficiently well structured 
causing no cache trashing. If using multiple core, this assumes that
no element of the {\tt x} vectors will be used by multiple
cores. \cite{Saule12} showed that there is very little cache trashing
in practice, but it showed that having elements of the vectors used by
multiple cores can have a significant impact on the performance
(growing with $n_v$).

On the other hand, in the worst case, every time the {\tt x} vector is
accessed, the value needs to be transfered from memory again. So in
total, there are as many transfers as the number of non-zeros in the
matrix. Note however that most architectures cannot read memory a single 
byte at a time. Instead, a minimum number of bytes,equal to the size o a cacheline, 
are transfered at once. 
%multiple bytes, with each chunk equal to the data do not allow to read the memory
%one byte at a time but always in a set of multiple bytes. On most
%cached architecture, that 
%the minimum size corresponds to the size $cl$ of
%a cacheline (typically 64 bytes). 
When there are multiple
vectors, each non-zero element uses $n_v$ consecutive entries. The
worst case number of bytes read and transfered are 
$$B_{rT} = s_M + n_c cl \ceil{\frac{n_vb_x}{cl}} $$ 
$$B_T = n_v n_m b_x n_r + n_r n_z \left ( b_i + b_x n_m +  cl \ceil{\frac{n_vb_x}{cl}} \right)$$

@@

In the special case $n_v=n_m=1$ in single precision
$$I_b = \frac{1}{4} \frac{ n_z }{ n_z + 1} \approx \frac{1}{4}$$
%$$I_w = \frac{n_z }{2  + 36  n_z} \approx \frac{1}{36}$$
$$I_w = \frac{n_z }{2  + (4+cl/2)  n_z} \approx \frac{1}{4+cl/2}$$

In the special case $n_v=n_m=4$ in single precision
$$I_b = \frac{32 n_z }{ n_z 20 + 80} \approx \frac{32}{20} = \frac{8}{5} $$
%$$I_w = \frac{32 n_z }{64 + n_z 84} \approx \frac{32}{84} = \frac{8}{21} $$
$$I_w = \frac{32 n_z }{64 + (20+cl) n_z} \approx \frac{32}{84} = \frac{8}{21} $$
The above formulas have been applied to MIC, which has a cache size $cl=64$ bytes.

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|r|}
    \hline
                           & v1m1  & v1m4  & v4m1   & v4m4  \\
    \hline
    Best Single Precision  & 36.36 & 58.18 & 133.33 & 213.33\\
    Worst Single Precision &  4.15 & 14.20 &  16.55 &  55.81\\
    Best Double Precision  & 24.00 & 32.21 &  85.71 & 117.07\\
    Worst Double Precision &  3.93 & 11.88 &  15.58 &  46.15\\
    \hline
  \end{tabular}
  \caption{Estimation of performance in GFlop/s at 150GB/s for a
    matrix of $n_z = 32$}
\end{table}

\begin{figure*}
  \centering 
  \subfigure[Absolute performance ]{\includegraphics[width=.49\linewidth]{figures/gflops_peak.pdf}\label{fig:gflops_peak_perf}}
%
  \subfigure[Relative performance to base case using one matrix and one
    vector.]{\includegraphics[width=.49\linewidth]{figures/speedup_wrt_base.pdf}\label{fig:speedup}}

  \caption{Estimation of the reachable performance one can
      achieve using a device with a memory to computational units of
      150GB/s when varying the number of vectors and matrices}
  \label{fig:perf_predict}
\end{figure*}

\todo{conclude on the evaluation using the figure/table. Note that we
  base prediction on 150GB/s which is not the peak of saule12. Note
  that if you are in best in v1m1 you might go to bad v4m4, but if you
  are in bad v1m1, you will essentially stay there.}

@@

\begin{figure*}
\centering
  \subfigure[
    Entries with "cpp" denote cases where
    coding was performed without MIC vector
    instructions.]{\includegraphics[width=.40\linewidth]{figures/test1_readwrite.pdf}\label{fig:band_rw}}
%
  \hspace{.3in}
%
  \subfigure[
    The speed on the CPU matches the speed with vector instructions.
    All memory is aligned on 32
    bytes.]{\includegraphics[width=.40\linewidth]{figures/host_test1_readwrite_no_temporal_hint.pdf}\label{fig:read_write}}

@@

\begin{figure*}
\centering
  \subfigure[ xxx ] 
  {\includegraphics[width=.20\linewidth]{figures/test1_gather.pdf}\label{fig:band_gather}} \\

  \subfigure[The greater
    speed of the cpp version is obtained throught the use of
    \ttt{\#Ivdep} {\em and\/} \ttt{\_\_assumed\_aligned}. All memory
    is aligned on 64
    bytes.]{\includegraphics[width=.20\linewidth]{figures/test3_gather.pdf}\label{fig:band_gather_ivdep}}
\end{figure*}
%

@@

\begin{figure*}
  \centering
  \subfigure[Performance of $y=Ax$ on the MIC. Squares: base 1/1 case,
    solid circles: 4/4 case implemented in C++, solid triangles: 4/4
    case implemented with MIC vector instructions. $-O3$ compilation
    options. Each symbol has two cases: with (red) and without blue) 
    with and without Reverse Cuthill-McKee. The grid is $96^3$.]
    {\includegraphics[width=.40\linewidth]{figures/mic_performance_nb_threads.pdf}\label{fig:perf_mic}}
%
\hspace{.4in}
%
  \subfigure[Performance on the host on a sparse matrix with $64^3$ rows. Manual
    vectorization brings significant improvement there as
    well.]{\includegraphics[width=.40\linewidth]{figures/64x64x64.pdf}\label{fig:perf_host}}
  \caption{Vectorization matters in practice. Differences in BW
    translate in different performance.\todo{this two plots should ``correlate'' }}
\end{figure*}

@@

%\begin{figure}
%  \centering
%  \includegraphics[width=.45\textwidth]{figures/plot_for_natasha_ncar_end_of_year.pdf}
%
%  \caption{Performance on a MIC of derivative computation for RBFFD
%    with 884,736 nodes distributed quasi-randomly in a cube. Single
%    derivative of a single function (base case, triangle) and four
%    derivatives of four functions (multi case, circles). No bandwidth
%    reduction (blue) and Reverse Cuthill-McKee (red).\todo{should we have a similar one for the host?}}
%\end{figure}
%
\begin{figure*}
\centering
\subfigure[Best case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_best-crop.pdf}} %\label{fig:ratio_best}}
\subfigure[Worst case.]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_worst-crop.pdf}}%\label{fig:ratio_worst}}
\subfigure[Worst case (No cacheline effects)]{\includegraphics[width=.33\linewidth]{figures/flops_to_bytes_no_cache-crop.pdf}}%\label{fig:ratio_nocache}}
\caption{Ratio of flops to bytes; (a) simple count of flops and bytes that require transfer from memory;(b) best and (c) worst
case taking into account that bringing one element of $x$ into memory brings along the remainder of the cache line.}
\label{fig:ratio_bytes_flops}
\end{figure*}
